---
title: "STAT 345 Final Project - Work(out) Smarter"
author: "Aaron Barton, Chase Tenner, Sean Hooyman, Lucas Zasada"
date: "2024-05-4"
output: word_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries
```{r message=FALSE, include=FALSE}
library(readxl)
library(tidyverse)
library(ggplot2)
library(shinydashboard)
library(shiny)
library(lubridate)
library(fmsb)
library(stringr)
```

To upload new data, the coach would need to import the data and change the name to the new file name in the R code. They should keep variable names consistent to these files so everything works.
```{r include=FALSE}
# Import data
data_updated = read_excel("Updated-Compiled_Data_Example.xlsx") 
report = read.csv("Updated-RPE Report.csv")
```

Data cleaning and manipulation
```{r include=FALSE}
# Omit NA values
data_updated = na.omit(data_updated)
report = na.omit(report)

#Summary for each activity
activity_summary = report %>% select(-Date, -ID) %>% group_by(Activity.Type) %>% summarize(across(everything(), mean))

# Remove categorical variables and summarize by player in regular dataset
summarized_data = data_updated %>% select(-Test_Date, -TestNumber) %>% group_by(ID) %>% summarize(across(everything(), mean))

# Remove categorical variables and summarize by player in report dataset
summarized_report = report %>% select(-Date, -Activity.Type) %>% group_by(ID) %>% summarize(across(everything(), mean))

# Summarized data by players with complete data
joined_data = inner_join(summarized_data, summarized_report, by = "ID")
# summary(joined_data)

# bins weeks together
report_weeks = report %>% mutate(Date_1 = as.Date(Date, format = "%m/%d/%Y")) %>% mutate(Week = floor_date(Date_1, unit = "week"))

# summarizes by week and retains activity type and player ID
summarized_activity_weeks = aggregate(sRPE ~ Week + Activity.Type, data = report_weeks, FUN = mean)
summarized_player_weeks = aggregate(sRPE ~ Week + ID, data = report_weeks, FUN = mean)

# averages the teams sRPE by week and combines it with the player's data
summarized_team_weeks = summarized_player_weeks %>% group_by(Week) %>% summarize(weeklyTeamAve = mean(sRPE))
summarized_player_weeks = inner_join(summarized_player_weeks, summarized_team_weeks, by = "Week")
```



Start by running all code and open the R with dashboard. Once opened, you can navigate between tabs by clicking on them depending on what metric you want to look at. Each tab has a drop down where you can search or select the option you want to be displayed in the graphic.

first tab: weekly boxplots for the entire team filtered by the type of activity

second tab: individual player analysis
-metric card: shows the average sRPE for last week for the filtered player
-bar and line chart: displays the weekly average sRPE with a line to compare to the overall team average
-spider plot: allows players and coaches to monitor important metrics that are inversely related to exertion. The lower the metrics, the more evidence of exertion. The plot has 2 shapes. One showing the filtered players baseline, which is the players overall average for all recorded time. The other shape shows the players average for the last week (currently this is just simulated data with variance from the overall baseline average). The plot turns red if the metrics are too low, indicating exertion. The plot turns green if the metrics are within the accepted range of the baseline. The threshold can be changed in the code.

```{r echo=FALSE, message = FALSE}
# Dashboard code
# Define UI

ui <- dashboardPage(
  dashboardHeader(title = "sRPE Dashboard"),
  dashboardSidebar(
    sidebarMenu(
      menuItem("Activity Summary", tabName = "activity_summary", icon = icon("chart-bar")),
      menuItem("Player Analysis", tabName = "player_analysis", icon = icon("user"))
    )
  ),
  dashboardBody(
    tabItems(
      #first tab for team analysis
      tabItem(tabName = "activity_summary",
              fluidRow(
                #dropdown for activity
                box(selectInput("activity_select", "Select Activity:", choices = unique(report$Activity.Type)), width = 12),
                box(plotOutput("activity_summary_plot"), width = 9)
              )
      ),
      #second tab for player analysis
      tabItem(tabName = "player_analysis",
              fluidRow(
                #dropdown for player ID
                box(selectInput("player_id", "Select Player ID:", choices = sort(unique(report$ID))), width = 6), 
                # metric card for average sRPE (last week)
                valueBoxOutput("average_srpe_box")
              ),
              fluidRow(
                # plot for players sRPE over time
                box(plotOutput("srpe_by_player_plot"), width = 6),
                # spider plot for monitoring exertion
                box(plotOutput("exertion_measures_by_player_plot"), width = 6)
              )
      )
    )
  )
)


# Define server logic
server <- function(input, output) {
  # Render histogram based on selected activity
  output$activity_summary_plot <- renderPlot({
    req(input$activity_select)
    filtered_report <- report_weeks %>% filter(Activity.Type %in% input$activity_select)
    ggplot(filtered_report, aes(x = Week, y = sRPE)) + geom_boxplot(aes(group = Week), fill = "royalblue") +
      labs(title = paste("Weekly Boxplots of sRPE for", paste(input$activity_select, collapse = ", "))) +
      xlab("Date") + ylab("sRPE")+
      theme(plot.title = element_text(face = "bold"))

  })

  # Render sRPE plot for the specified player ID
  output$srpe_by_player_plot <- renderPlot({
    req(input$player_id)
    # filter by player ID
    report_filtered <- summarized_player_weeks %>% filter(ID == as.numeric(input$player_id))
    ggplot(report_filtered, aes(x = Week, y = sRPE)) + geom_col(fill = "seagreen") + 
      geom_line(data = report_filtered, aes(x = Week, y = weeklyTeamAve, color = "firebrick"), size = 1) +
      labs(title = paste("Weekly Perceived Exertion Averages for Player ID", input$player_id)) +
      scale_color_manual(values = "firebrick", labels = "Weekly Team Average", name = "") +
      xlab("Date") + ylab("sRPE")+
      theme(plot.title = element_text(face = "bold"))

  })
  
  # Render exertion measures plot for the specified player ID
  output$exertion_measures_by_player_plot <- renderPlot({
    req(input$player_id)
    # select ID and relative exertion metrics
    variables <- c("ID", "Relative Force at Min Displacement", "IMTP", "MAS_Ave(mph)", "Fast_Reaction")
    spider_data <- joined_data %>% select(all_of(variables)) %>% select(-ID)
    # calculate min and max values of each variable for scaling the spider plot
    max_values <- apply(spider_data, 2, max)
    min_values <- apply(spider_data, 2, min)
    
    # get individual data for filtered player
    data_individual <- joined_data %>%
      select(all_of(variables)) %>%
      filter(ID == as.numeric(input$player_id)) %>%
      select(-ID)
  
    data_individual <- rbind(min_values, data_individual)
    data_individual <- rbind(max_values, data_individual)
    
    # make simulated data for new week
    new_data_individual <- data_individual %>%
      slice(3)
    normal_variation <- rnorm(n = ncol(new_data_individual), mean = 0, sd = 0.5)
    new_data_individual <- new_data_individual + normal_variation
    new_data_individual <- rbind(min_values, new_data_individual)
    new_data_individual <- rbind(max_values, new_data_individual)
    
    #scaling the shapes
    student_overall_avg <- data_individual %>%
      slice(3)
    new_weekly_avg <- new_data_individual %>%
      slice(3)
    new_data_individual <- rbind(new_data_individual, student_overall_avg)
    
    
    # Threshold is permissable range that new week average can go below the baseline before setting off red alert.
    threshold <- 0.01
    newcolor <- ifelse(any(new_weekly_avg > student_overall_avg - threshold), "green", "red")
    color <- c(newcolor, "grey")
    
    radarchart(new_data_individual,
               plty = 3,
               pcol = "#636363",
               plwd = 2,
               pfcol = scales::alpha(color, 0.8), # translucent colors
               cglty = 1,
               cglcol = "#3b3b3b",
               title = str_wrap("Monitoring Key Metrics of Exertion (Player Baseline Average vs. Average for Last Week)", width = 50),
               vlcex = 1)
    
    legend("bottomright", legend = c("Out of Safe Range", "In Safe Range"),
           fill = c("red", "green"), title = "Legend")
  })
  
  
  

 # Render average sRPE for the latest week for the selected player
  output$average_srpe_box <- renderInfoBox({
    
    req(input$player_id)
    # get the latest week average for the filtered player
    latest_week_avg <- summarized_player_weeks %>%
      filter(ID == as.numeric(input$player_id)) %>%
      arrange(desc(Week)) %>%
      slice(1) %>%
      pull(sRPE)  # Extracting the average sRPE for the latest week
    
valueBox(
    subtitle = "Last Week's Average sRPE", 
    value = latest_week_avg,
    color = "light-blue",
    width = 30)
    
    })

}

# Run the application
shinyApp(ui, server)

```

